Psyc 601 Multivariate Statistics Project

### 2013-08-26 writeups/proposal/proposal.tex
foo
Former-commit-id: e8e06245abfd0c0b47542c07258ee8dbc5937b43 [formerly 4ce8434aebd15fc50381302c2185c07d74c9e4e8]
Former-commit-id: e20d7ecf4e64580855d3b11817e36ea8b0a7232f
foo
Former-commit-id: ebd82cfcb57fb06af8059c0ca75e6fbbc3bfb82d [formerly b356864c73b6b70f2a7d0e5a6d73f5503ec7438d]
Former-commit-id: c554956320e46828e29582d53edba33dd6a8c4ad

To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
There are two likely reasons for this:
[1] Modeling user-chosen hashtags and building hashtag-recommendation systems has only explored over the past few years, and
[2] Twitter is a large database of human-generated content, spanning a wide range of topics, containing a sizeable set of user-generated hashtags, so hashtag prediction for Twitter is a relatively hard problem.
We are nonetheless curious if declarative memory retrieval models can be applied to this domain to generate useful hashtag predictions.
The domain shares many similarities with environments where hashtag-recommendation systems have already been deployed:
Google+'s microblogging service is essentially Twitter, but focused on a specific subset of users and topics.
The StackExchange sites are also built from a large database of human generated content, where users have chosen specific tags to associate with posts.
So we will evaluate how well an ACT-R-inspired Bayesian declarative memory model and a new vector-based model predict a user's chosen hashtag when composing a tweet on Twitter.

@@

Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example, can be represented as each word in the title and body of a post.
Context for the Twitter domain would be the words in a Tweet.
Association strenght between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likeihood that the chunk occured in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).

>>>>

Several hashtag recommendation models for Twitter have been developed recently.
One of the first models used a Bayesian co-occurrence statistical technique to predict the most likely hashtag associated with a post as a function of prior hashtag use and context. \parencite{Mazzia2009}
The model present by \textcite{Mazzia2009} is very similar to an ACT-R declarative retrieval model.
The main potential difference is that the global prior likelihood of a hashtag is computed without taking into account the specific user's past history.
Instead, the global prior is an overall average frequency of hashtag use.

Several other models have taken a tweet-centered approach to hashtag prediction, where suggested hashtags are collected from hashtags used from similar tweets.
\textcites{Li2011, Zangerle2011, Kywe2012} all store a content vector for each tweet, and then compute a word co-occurence-based similarity score between a composed tweet and the rest of the tweets in the database.
\textcites{Zangerle2011, Kwye2012} use the Apache Lucine infrastructure to compute the similarity score, while \textcite{Li2011} uses a custom similarity score derived from the WordNet database to compute similarites.
Regardless of the method to compute the similarity between tweets, each method collects the hashtags used from a set of the most similar tweets, then ranks the set, and presents the top 5-10 hashtags to the user.

This tweet-centered approach is quite different than the hashtag-centered approach presented in \textcite{Mazzia2009}.
With a hashtag-centered approach, a direct association is built between the words in a tweet and hashtags that occur in tweets.
With a tweet-centered approach, that association between words and hashtags is indirect:
The relationship is built between words in a tweet and a user, and then similar users are assumed to use similar hashtags.
One problem with the tweet-centered approach is that the storage size grows with the number of tweets, as a vector representatino of the content in each tweet has to be maintained to later compute tweet similarity.
The storage size for a hashtag-centered approach grows with the number of hashtags, but it seems reasonable that this will grow slower than the number of tweets and asymptote after collecting a large sample of tweets.
So from the perspective of efficient information storage and retrieval, it seems more likely that the hashtag-centered approach used in \textcite{Mazzia2009} will scale as Twitter grows in size and use.

\textcite{kwye2012} also customized their tag predictino model to the user's past hashtag use, and was one of the first to do so when developing a hashtag-prediction model for Twitter.
The set of recommended hashtags was the union of hashtags used in similar tweets and hashtags used by similar users.
This was done by storing a content vector of hshtag use for each user (user-centered approach) alongside the content vectors of hashtag use for each tweet (tweet-centered approach).
To generate recommended hashtags, the model would return the top-ranked hashtags from the hashtags used in 0-50 similar tweets and 0-4 similar users.
Prediction accuracy improved when the model generated recommendations based on hashtags in both similar tweets and the top few similar users compared to basing recommendations on similar tweets alone.

FIXME: \cite{Godin2013}
This paragraph will discuss the topics-based approach.

FIXME: Paragraph that no vector-based models have been attempted for Twitter.
Bayesian model can be improved by customizing predicitons based on past hashtag use.

FIXME: ditch this paragraph?
To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
There are two likely reasons for this:
[1] Modeling user-chosen hashtags and building hashtag-recommendation systems has only explored over the past few years, and
[2] Twitter is a large database of human-generated content, spanning a wide range of topics, containing a sizeable set of user-generated hashtags, so hashtag prediction for Twitter is a relatively hard problem.
We are nonetheless curious if declarative memory retrieval models can be applied to this domain to generate useful hashtag predictions.
The domain shares many similarities with environments where hashtag-recommendation systems have already been deployed:
Google+'s microblogging service is essentially Twitter, but focused on a specific subset of users and topics.
The StackExchange sites are also built from a large database of human generated content, where users have chosen specific tags to associate with posts.
So we will evaluate how well an ACT-R-inspired Bayesian declarative memory model and a new vector-based model predict a user's chosen hashtag when composing a tweet on Twitter.

@@

Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example, can be represented as each word in the title and body of a post.
Context for the Twitter domain would be the words in a Tweet.
Association strenght between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likeihood that the chunk occured in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.
